Adapting the MVDR to a GPU for active sonar systems of different sizes.

### 2013-01-22 ASA2013/asa2013.tex
la2pdf commit - 2013:01:22	11:07:12	(Tue22.Jan)
la2pdf commit - 2013:01:22	11:39:43	(Tue22.Jan)

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1. \label{lca_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
Note that (\ref{mvdr_ref}) and (\ref{lca_ref}) is essentially the same criterion.

@@

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1. \label{lca_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
Note that (\ref{mvdr_ref}) and (\ref{lca_ref}) is essentially the same criterion.

@@

The apparent solution is make sure that each thread primarily access the shared memory (L1 cache) and registers, but these are scarce resources. 
\begin{table}[b]\centering%\normalsize
\begin{tabular}[c]{l r r r@{:}  l}\hline
\rowcolor{tabBlue} & \multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{arithmetic}$} & \multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{memory}$} & \multicolumn{2}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{memory}$/BW$_\text{arithmetic}$} \\\hline
Arithmetic throughput & 1.03 Tflop/s & & \hspace{30pt} 1 & 30 \\
Global memory & & 36 Gfloats/s & \hspace{30pt} 1 & 30 \\
Shared memory & & 257 Gfloats/s & 1 & 4 \\
Registers & & $>$1.5 Tfloats/s & $>$3 & 2~\cite{Vasilyy}
\end{tabular}
\caption{Nvidia Quadro 6000; Memory bandwidths (BW) compared to the peak arithmetic bandwidth of 1.03 TB/s.}\label{tabbandwidth}
\end{table}

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=0.7\linewidth]{gfx/mvdr_implementation.svg}%
\caption{MVDR implementation.}\label{mvdr_implementation}
\end{figure}

>>>>

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1\label{lca_def}. 
% \vec w[n] = \frac{\Ri[n]\1}{\1\T\Ri[n]\1},\label{weights}
\end{gather}
Note that (\ref{mvdr_ref}) and (\ref{lca_ref}) is essentially the same criterion.

@@

The Low Complexity Adaptive (LCA) beamformer is another take on the optimization problem in (\ref{mvdr_def}). Instead of calculating the weights like in (\ref{weights}), it simply computes the beamformer output for each window in a collection of such, and finally applies the one that best fulfils the given optimization criterion. If we let $\w_p$ be the $p$'th of a total of $P$ windows, then this can be formulated as:
\begin{gather}
\argmin{p}\ E\Big\{\big|\w_p\H\x[n]\big|^2\Big\} \qquad\text{subject to}\qquad \w_p\H\1 = 1. \label{lca_def}
\end{gather}
Note the similarity between (\ref{mvdr_def}) and (\ref{lca_def}). The LCA and MVDR uses essentially the same optimization criterion, but whereas MVDR searches a continuous solution space for the best set of weights to apply, the LCA beamformer only have $P$ solutions to select from. However, once the LCA have 10-20 well designed windows to choose from its performance quickly converge towards the MVDR beamformer's performance. The difference in computational complexity, however, is

@@

The solution is make sure that each thread primarily access the shared memory (L1 cache) and registers. Unfortunately, these are scarce resources and 
\begin{table}[b]\centering%\normalsize
\begin{tabular}[c]{l r r r@{:}  l}\hline
\rowcolor{tabBlue} & \multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{arithmetic}$} & \multicolumn{1}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{memory}$} & \multicolumn{2}{>{\columncolor{tabBlue}}c}{\bf BW$_\text{memory}$/BW$_\text{arithmetic}$} \\\hline
Arithmetic throughput & 1.03 Tflop/s & & \hspace{30pt} 1 & 30 \\
Global memory & & 36 Gfloats/s & \hspace{30pt} 1 & 30 \\
Shared memory & & 257 Gfloats/s & 1 & 4 \\
Registers & & $>$1.5 Tfloats/s & $>$3 & 2~\cite{Vasilyy}
\end{tabular}
\caption{Nvidia Quadro 6000; Memory bandwidths (BW) compared to the peak arithmetic bandwidth of 1.03 Tflop/s.}\label{tabbandwidth}
\end{table}

\begin{figure}[t]\centering%
\graphicsAI[drawing,width=\linewidth]{gfx/mvdr_implementation.svg}%
\caption{MVDR implementation.}\label{mvdr_implementation}
\end{figure}
